\chapter{Get a tracefile}\label{cha:GetATrace}

In order to use the folding framework it is imperative to use tracefiles with performance metrics gathered using instrumentation and sampling.
This chapter provides some \TRACE configuration skeleton files that the analyst may use so as to instruct \TRACE to generate tracefiles ready to be processed through the \FOLDING mechanism.
Note that these configuration files are extremely simplistic and only focus on setting an appropriate set of performance counters and ensure that the sampling mechanism is enabled (in these examples with a sampling period of 25~milliseconds).
For any further information with respect to the \TRACE configuration, please refer to the \TRACE user guide available at the BSC website. 

Additionally, the folding mechanism is able to combine several performance models and generate summarized results that simplify understanding the behavior of the node-level performance.
Since these performance models are heavily-tighted with the performance counters available on each processor architecture and family, the following sections provide Extrae XML configuration files ready to use on several architectures.
Note also, that an XML configuration file is provided in subsection~\ref{subsec:OtherArchitectures} for other architectures not listed here.
Since each architecture has different characteristics, the user may need to tune the XML presented there to make sure that all the list performance counters are gathered appropriately.

\section{XML configurations}

\newpage

\subsection{For Intel SandyBridge processors}

\input{XML/extrae-intel-sandybridge.xml}

\newpage

\subsection{For Intel Nehalem processors}

\input{XML/extrae-intel-nehalem.xml}

\newpage

\subsection{For the remaining architectures}\label{subsec:OtherArchitectures}

\input{XML/extrae-basic.xml}

